---
title: "Regression Models Course Project"
author: "Mohit Shukla"
date: "Sunday, May 24, 2015"
output: pdf_document
---

##### Summary    

This report includes an analysis of mtcars data set which is a collection of cars, main interest of  exploration is to understand the relationship between a set of variables and miles per gallon (MPG) (outcome). Main focus of the analysis is the two following aspects:

* "Is an automatic or manual transmission better for MPG"
* "Quantify the MPG difference between automatic and manual transmissions"

This analysis may follow many models depending upon the number of covariates being included. Hence A model comparison test is performed to test the relative reliablility of different models. A number of exploratory analysis tasks have been performed to check the validity of models.

The analysis shwoed that all covariates produces more reliable model than single covariate model. Since, a motar car has so many influencing factors other than transmission, so this result seems logical.
From including two covariates to including all the covariates, results of the model didn't show much difference. Hence **all covariate model has been selected** over all the others assuming that it represents all the other models and dominates the single covariate model.


**Analysis of mpg based on factor variable am** 

```{r}
data(mtcars)
data <-  mtcars[,c("mpg","am")]
data[,"am"] <-  as.factor(data[,"am"])
attach(data)
summary(mpg[am == 1])
summary(mpg[am == 0])
fit <- lm(mpg~am)
summary(fit)$coef

```
Above analysis shows dominance of am==1 i.e. Manual transmission. Since the data were coded as 0/1, the coefficient is directly interpreted as the manual (1) - auto (0) difference. In other words, manual on average were 7.245 points greater than auto, a significant difference.

**Analysis of mpg based on all the covariates**

```{r}
data(mtcars)
model <- lm(mpg~., data=mtcars)
summary(model)$coef
```
The above model shows that manual transmission on average is `r (summary(model)$coef["am",])[1]` points better than the auto transmission.
Now this is serious **conflict** between sigle covariate model and all covariate model.
To decide between these two and all the other available models a model selection test is performed below.

**Model Selection**

```{r}

data(mtcars) 
fit10 <- lm(mpg~ am, data=mtcars)
fit9 <- update(fit10, mpg~  cyl + am, data=mtcars)
fit8 <- update(fit10, mpg~  cyl + disp + am, data=mtcars)
fit7 <- update(fit10, mpg~  cyl + disp + hp + am, data=mtcars)
fit6 <- update(fit10, mpg~  cyl + disp + hp + drat + am, data=mtcars)
fit5 <- update(fit10, mpg~  cyl + disp + hp + drat + wt + am, data=mtcars)
fit4 <- update(fit10, mpg~  cyl + disp + hp + drat + wt + qsec + am , data=mtcars)
fit3 <- update(fit10, mpg~  cyl + disp + hp + drat + wt + qsec + vs+ am , data=mtcars)
fit2 <- update(fit10, mpg~  cyl + disp + hp + drat + wt + qsec + vs+ am + gear, data=mtcars)
fit1 <- update(fit10, mpg~ cyl + disp + hp + drat + wt + qsec + vs+ am + gear + carb, data=mtcars)
anova(fit10,fit9,fit8,fit7,fit6,fit5,fit4,fit3,fit2,fit1)
anova(fit10,fit1)
```
The above analysis shows (by looking at Pr>F values )that all the models from model1 to model9 are better off than model10 which includes only one covariate.
Second result shows the dominance of **all covariates model** over singel covariate model.  
Hence it is safe to go with model that includes all the covariates i.e. **fit1** model.

**Result**

* coefficients for "am" are: `r summary(fit1)$coef["am",]`
* Manual Transmission (am==1) is `r (summary(fit1)$coef["am",])[1]` points better than Automatic Transmission i.e. (am==0)           

**Appendix**

```{r echo=FALSE}
#plot1 <- pairs(mtcars, panel=panel.smooth, main="Plot1: mtcars data", col=3+(mtcars$am==1))
#mtcars[,9] <- as.factor(mtcars[,9])
par(mfrow=c(1,1))
plot2 <- boxplot(list(manual=mpg[am==1], automatic=mpg[am==0]), main = "Plot2: box plot for mpg~am" )
#plot3 <- plot(am, mpg, main="Plot3: am and mpg", pch=19)
par(mfrow=c(2,2))
plot4 <- plot(fit, main = "residual plots for mpg~am")
par(mfrow=c(2,2))
plot5 <- plot(fit1, main = "residual plots for mpg~.")
```







